---
title: "Import and Tidy"
author: "David Henderson"
date: 2018-02-01
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

#Introduction

On this page I show how to import and tidy the three years of open data

##Aims
* Import data into R
* Tidy and coerce variables
* Save data as .RData objects

##Packages
Load required packages
```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(forcats)
library(cowplot)
```

#Import Data

I've downloaded the data files (see home page) into the "data" project folder and I'll imort them into R now. If you are copying the code you will need to set the file paths relative to your system. 

```{r import, message=FALSE, warning=FALSE, echo=FALSE}
#chunk not rendered. Actual import so filepath isn't shown on website
soc_care10 <- read_csv("C://GitHub/Scottish_Social_Care_Open_Data/data/public_release_sc_2010.csv")
soc_care11 <- read_csv("C://GitHub/Scottish_Social_Care_Open_Data/data/public_release_sc_2011.csv")
soc_care12 <- read_csv("C://GitHub/Scottish_Social_Care_Open_Data/data/public_release_sc_2012.csv")
```

```{r import_dummy, eval=FALSE}
soc_care10 <- read_csv("...~/data/public_release_sc_2010.csv")
soc_care11 <- read_csv("...~/data/public_release_sc_2011.csv")
soc_care12 <- read_csv("...~/data/public_release_sc_2012.csv")
```

#Tidy data

Let's start by looking at each set of data

##Glimpse

```{r glimpse10}
glimpse(soc_care10)
glimpse(soc_care11)
glimpse(soc_care12)
```

We can see in 2010 there are `r nrow(soc_care10)` observations of `r ncol(soc_care10)` variables. 2011 has `r nrow(soc_care11)` observations of `r ncol(soc_care11)` variables and 2010 has `r nrow(soc_care12)` observations of `r ncol(soc_care12)` variables. 

The reason for the discrepancies in observations is that 2011 and 2012 data contains information on individuals who received not just home care, but also telecare which almost doubles the number of people included. 

3 telecare variables are added in 2011. In 2012 these 3 variables are present plus an extra variable `newhours`. There are also 3 unneeded variables in 2012; `error`, `error1`, and `HCFlag`. The first two are self-explanatory. The third is a duplicate of the `HCclient` variable. I'll drop these now before we start to tidy. 

```{r drop_12_unneded}
soc_care12 %<>% 
  select(-error, -error1, -HCFlag) 
```

With the exception of`clientID` in each year, every variable needs coerced. The majority of these variables should be factors. I could re-import the data and try `stringsasFactor == TRUE`. I'm not so sure this would work and, anyhow, I want to explicitly name the labels for the factors. 

1st of all though, we'll sort out the variable names. 

#New names

```{r tolower}
#different number of coulmns in each year so requires seperate commands

names(soc_care10) <- c("council", "client_id", "age_grp", "client_grp", "gender", "meals",
                       "la_hrs", "pri_hrs", "vol_hrs", "pc_hrs", "total_hrs", "hc_client",
                       "living_arr", "multi_staff", "housing_type", "laundry", "shopping",
                       "housing_supp", "iorn_score")

names(soc_care11) <- c("council", "client_id", "age_grp", "client_grp", "gender", "meals",
                       "la_hrs", "pri_hrs", "vol_hrs", "pc_hrs", "total_hrs", "hc_client",
                       "comm_alarm", "other_telecare", "alarm_and_tele", "living_arr", 
                       "multi_staff", "housing_type", "laundry", "shopping",
                       "housing_supp", "iorn_score")

names(soc_care12) <- c("council", "client_id", "age_grp", "client_grp", "gender", "meals",
                       "new_hrs", "la_hrs", "pri_hrs", "vol_hrs", "pc_hrs", "total_hrs",
                       "hc_client", "comm_alarm", "other_telecare", "alarm_and_tele", 
                       "living_arr", "multi_staff", "housing_type", "laundry", "shopping",
                       "housing_supp", "iorn_score")
```

#Coerce variables.

I'll coerce each in order here. I'm being explicit so I can set the labels for each factor level. All labels and levels identified from the [metadata file](http://www.gov.scot/Resource/0040/00409164.pdf).

##Local Authority Code

```{r factorise_LAcode}
la_levels <- c("100", "110", "120", "130", "150", "170","180","190","200", "210", "220",
              "230", "235", "240", "250", "260","270", "280", "290", "300", "310", "320",
              "330", "340","350", "355", "360", "370", "380", "390", "395", "400")
la_labels <- c("Aberdeen City", "Aberdeenshire", "Angus", "Argyll & Bute", "Clackmannanshire",
              "Dumfries & Galloway", "Dundee City", "East Ayrshire", 
              "East Dunbartonshire","East Lothian", "East Renfrewshire",
              "Edinburgh, City of", "Eilean Siar", "Falkirk", "Fife", "Glasgow City",
              "Highland","Inverclyde", "Midlothian", "Moray", "North Ayrshire",
              "North Lanarkshire", "Orkney Islands","Perth & Kinross", "Renfrewshire",
              "Scottish Borders", "Shetland Islands","South Ayrshire", "South Lanarkshire",
              "Stirling","West Dunbartonshire","West Lothian")

soc_care10$council <- factor(soc_care10$council,
                            levels = la_levels,
                            labels = la_labels)

soc_care11$council <- factor(soc_care11$council,
                            levels = la_levels,
                            labels = la_labels)

soc_care12$council <- factor(soc_care12$council,
                            levels = la_levels,
                            labels = la_labels)
```

Now I want to count each level of the factors as a sanity check. Do they look roughly like they should do? (I'll explore in more detail in another file). 


```{r count_la_level}
fct_count(soc_care10$council)  #Count factor levels as diagnostic check
fct_count(soc_care11$council)
fct_count(soc_care12$council)

#remove unneeded object from memory
rm(list = c("la_labels", "la_levels"))
```

##Age Group

`age_grp` has been imported as an integer - factorise. I'll also convert missing data level to `NA` 

```{r factorise_AgeGRP}
age_grp_levels = c("1", "2", "3", "4", "5")
age_grp_labels <- c("18-<65", "65-<75", "75-<85", "85+", "Unknown")

#2010

soc_care10$age_grp <- factor(soc_care10$age_grp,
                            levels = age_grp_levels,
                            labels = age_grp_labels)
soc_care10$age_grp[soc_care10$age_grp == "Unknown"] <- NA #Convert Missing data to NA
soc_care10$age_grp <- fct_drop(soc_care10$age_grp, only = "Unknown") #Drop unknown level


#2011

soc_care11$age_grp <- factor(soc_care11$age_grp,
                            levels = age_grp_levels,
                            labels = age_grp_labels)
soc_care11$age_grp[soc_care11$age_grp == "Unknown"] <- NA #Convert missing data to NA
soc_care11$age_grp <- fct_drop(soc_care11$age_grp, only = "Unknown")

#2012

soc_care12$age_grp <- factor(soc_care12$age_grp,
                            levels = age_grp_levels,
                            labels = age_grp_labels)
soc_care12$age_grp[soc_care12$age_grp == "Unknown"] <- NA #Convert Missing data to NA
soc_care12$age_grp <- fct_drop(soc_care12$age_grp, only = "Unknown")
```

And another sanity check

```{r count_age}
fct_count(soc_care10$age_grp) #Count factor levels as diagnostic check
fct_count(soc_care11$age_grp)
fct_count(soc_care12$age_grp)

#remove unneeded objects from memory
rm(list = c("age_grp_labels", "age_grp_levels"))

```

## Client Group

`client_grp` came in as an integer for 2010 but as a character for 2011 and 2012. I will recode them all to factors and add the labels. There are *two* missing data levels here! Both to `NA`

```{r factorise_ClientGRP}
client_grp_levels <- c("1", "2", "3", "4", "5", "99", ".")
client_grp_labels <- c("Dementia & Mental Health", "Learning Disability", 
                     "Physical Disability", "Infirmity due to Age","Other",
                     "Not Known", "Blank")

#2010

soc_care10$client_grp <- factor(soc_care10$client_grp,
                               levels = client_grp_levels,
                               labels = client_grp_labels)

soc_care10$client_grp <- fct_collapse(soc_care10$client_grp,
                                      missing = c("Not Known", "Blank"))

soc_care10$client_grp[soc_care10$client_grp == "missing"] <- NA #Covert missing data to NA
soc_care10$client_grp <- fct_drop(soc_care10$client_grp, only = "missing") #drop missing level


#2011

soc_care11$client_grp <- factor(soc_care11$client_grp,
                               levels = client_grp_levels,
                               labels = client_grp_labels)

soc_care11$client_grp <- fct_collapse(soc_care11$client_grp,
                                      missing = c("Not Known", "Blank"))

soc_care11$client_grp[soc_care11$client_grp == "missing"] <- NA #Covert missing data to NA
soc_care11$client_grp <- fct_drop(soc_care11$client_grp, only = "missing") #drop missing level


#2012
soc_care12$client_grp <- factor(soc_care12$client_grp,
                               levels = client_grp_levels,
                               labels = client_grp_labels)
soc_care12$client_grp <- fct_collapse(soc_care12$client_grp,
                                      missing = c("Not Known", "Blank"))

soc_care12$client_grp[soc_care12$client_grp == "missing"] <- NA #Covert missing data to NA
soc_care12$client_grp <- fct_drop(soc_care12$client_grp, only = "missing") #drop missing level
```

Sanity check on client group

```{r client_check}
fct_count(soc_care10$client_grp) #Count for diagnostic purposes
fct_count(soc_care11$client_grp)
fct_count(soc_care12$client_grp)

#remove unneeded objs
rm(list = c("client_grp_labels", "client_grp_levels"))

```


##Gender

`gender` again imported as an integer in 2010 but as a charcterin other 2 years. Again will recode all years to add labels. 

```{r factorise_GenderISO}
#2010
soc_care10$gender <- factor(soc_care10$gender,
                               levels = c("1", "2", "."),
                               labels = c("Male", "Female", "Blank"))
soc_care10$gender[soc_care10$gender == "Blank"] <- NA #Convert missing to NA level
soc_care10$gender <- fct_drop(soc_care10$gender, only = "Blank") #and drop empty level

#2011
soc_care11$gender <- factor(soc_care11$gender,
                               levels = c("1", "2", "."),
                               labels = c("Male", "Female", "Blank"))
soc_care11$gender[soc_care11$gender == "Blank"] <- NA #Convert missing to NA level
soc_care11$gender <- fct_drop(soc_care11$gender, only = "Blank") #and drop empty level

#2012
soc_care12$gender <- factor(soc_care12$gender,
                               levels = c("1", "2", "."),
                               labels = c("Male", "Female", "Blank"))
soc_care12$gender[soc_care12$gender == "Blank"] <- NA #Convert missing to NA level
soc_care12$gender <- fct_drop(soc_care12$gender, only = "Blank") #and drop empty level
```

Gender sanity check
```{r gender_check}
fct_count(soc_care10$gender)
fct_count(soc_care11$gender)
fct_count(soc_care12$gender)
```


##Meals

Factorise meals - dichotomous variable. **The 2011 and 2012 .csv files only seem to have values for meals meaning we have to presume empty values are "no meals", rather than missing data.** Not too sure about this but will investgate further in the analysis. For now the data stays as missing and will be computed as NA 

```{r factorise_meals}
soc_care10$meals <- factor(soc_care10$meals,
                           levels = c("0", "1"),
                           labels = c("No meals", "Meals")) 


soc_care11$meals <- factor(soc_care11$meals,
                           levels = c("0", "1"),
                           labels = c("No meals", "Meals"))


soc_care12$meals <- factor(soc_care12$meals,
                           levels = c("0", "1"),
                           labels = c("No meals", "Meals")) 

#Slightly different coding from metadata to make Graph labels more readable
```

Meals sanity check
``` {r meals_check}
fct_count(soc_care10$meals)
fct_count(soc_care11$meals)
fct_count(soc_care12$meals)
```

The Scottish Government report of the Social Care Survey states that Local Authorities found it difficult to collect and return data on meals. We can see that when visualising returns by Local Authority - I'm using 2012 as an example here:- 

```{r meals_by_LA, fig.height=12, fig.width=16}
library(ggplot2)
soc_care12 %>%
  ggplot(aes(x = meals)) +
  geom_bar(aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~ council)
```

Missing data does not seem to be at random here suggesting there is more of a structural issue with reporting of meals data. 

I'll drop this variable from the datasets

```{r drop_meals}
soc_care10 %<>% select(-meals)

soc_care11 %<>% select(-meals)

soc_care12 %<>% select(-meals)
```






## Session information

<!-- Insert the session information into the document -->
```{r session-info, cache=FALSE}
```
