---
title: "Functions"
author: "David Henderson"
date: 2018-02-06
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

# A list of my own created functions. 

These functions are stored in the "code" project folder and are called from there for any analysis. This page provides a quick reference for the functions

##factor_plot()

`factor_plot()` is function which has two arguments `df` and `group`. The former refers to the dataframe to use, the latter the variable to be plotted. 

The dataframe is grouped by the `group` variable, factor levels are counted and propotions calculated. This is then fed into a ggplot which plots the values using `geom_lollipop()` from the `ggalt` package. Many examples in [Exploratory Data Analysis - 2010 data](02_2010_eda.html)

```{r factor_plot, eval=FALSE}
factor_plot <- function(df, group){

  group <- enquo(group)

  x <-
    df %>%
    group_by(!!group) %>%
    summarise(N = n()) %>%
    mutate(pct = round((N/sum(N)) * 100, 1))

  ggplot(x, aes_(substitute(group), substitute(N), colour = substitute(group))) +
    geom_lollipop(point.size = 3) +
    geom_text(aes(label=paste0(pct,"%")), size=3, hjust = -0.3, vjust = -0.2) +
    scale_fill_manual(values = ubdc_palette) +
    theme(legend.position = "none") +
    labs(y = "Count",
         title = "Number of individuals receiving social care")

}
```

##x_45

This is more of a ggplot theme than a function. Often, depending on the plot, I need to tilt the label on the x-axis so avoid them overlapping. It is one line of code but I honestly have written it so many times it bugs me! This is an easy wrapper to save my poor fingers

```{r x_45, eval=FALSE}
x_45 <- theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8))

```
wflow
## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
